Incentivizing Exploration with Unbiased History

### 2018-10-31 model.tex
Another revision of prelims.

We study the multi-armed bandit problem in a social learning context, in which a principal faces a sequence of $T$ myopic agents. There is a set $\A$ of $K$ possible actions, a.k.a. \emph{arms}. At each round $t\in [T]$, a new agent $t$ arrives, receives a message $m_t$ from the principal, chooses an arm $a_t\in \A$, and collects a reward $r_t\in \{0,1\}$ that is immediately observed by the principal. The reward from pulling an arm $a\in \A$ is drawn independently from Bernoulli distribution $\cD_a$ with an unknown mean $\mu_a$. An agent does not observe anything from the previous rounds, other than the message $m_t$. The problem instance is defined by (known) parameters $K,T$ and the (unknown) tuple of mean rewards, $(\mu_a:\,a\in\A)$. We are interested in \emph{regret}, defined as
\[\textstyle 
  \reg(T)
  %= \reg(a_1, \ldots , a_T) 
  = T \max_{a\in \A} \mu_a -
  \sum_{t\in [T]} \E[\mu_{a_t}].
\]
The principal chooses messages $m_t$ according to an online algorithm called \emph{disclosure policy}, with a goal to minimize regret. We assume that mean rewards are bounded away from $0$ and $1$, to ensure sufficient entropy in rewards. For concreteness, we posit that
    $\mu_a\in [\tfrac13,\tfrac23]$.

\xhdr{Unbiased subhistories.}  
The \emph{subhistory} for a subset of rounds $S\subset [T]$, and the corresponding \emph{anonymized subhistory} are defined, respectively, as
\[ \SubH{S} = \{\; (s,a_s,r_s):\;s\in S \;\}
\quad\text{and}\quad
   \AnonSubH{S} = \{\; (a_s,r_s):\;s\in S \}.
\]
Accordingly, $\SubH{[t-1]}$ is called the \emph{full history} at time $t$.

@@

Further, we assume that disclosure policies are \emph{transitive}, in the following sense:
\[ t'\in S_t \Rightarrow S_{t'}\subset S_t
    \qquad \text{for all rounds $t,t'\in [T]$}. \]
In words, if agent observes the subhistory for some previous agent, then she observes the entire message for that agent.

A transitive unbiased-history policy $\pi$ can be represented as an undirected graph $G_\pi$, where nodes correspond to rounds, and any two rounds $t'<t$ are connected if and only if $t'\in S_t$ and there is no intermediate round $t''$ with
    $t'\in S_{t''}$ and $t''\in S_t$.
This graph is henceforth called the \emph{information flow graph} of policy $\pi$, or \emph{info-graph} for short.

\xhdr{Agents' behavior.} Let us define agents' behavior in response to an unbiased-history policy. We posit that each agent $t$ uses its observed subhistory $m_t$ to form a reward estimate $\hat{\mu}_{t,a} \in [0,1]$ for each arm $a\in \A$, and chooses an arm with a maximal estimator. (Ties are broken according to an arbitrary rule that is the same for all agents.)
We make some mild assumptions:

@@

\begin{assumption}\label{ass:embehave}
Reward estimates satisfy the following assumptions:
\begin{itemize}
\item[(a)] Reward estimates are close to empirical averages. Let $N_{t,a}$ and $\bar{\mu}_{t,a}$ denote the number of pulls and the empirical mean reward of arm $a$ in subhistory $m_t$. Then for some absolute constant $N_0\in \N$ and $C_0=\tfrac{1}{16}$ it holds that
\[
\forall t\in[T],\, a\in\A \qquad
N_{t,a} \geq N_0 \quad\Rightarrow\quad
    \left|\hat{\mu}^t_a - \bar{\mu}^t_a \right| <
		\frac{C_0}{\sqrt{N_{t,a}}},
\]
\item[(b)] Reward estimates have a warm-start property:
\[
\forall t\in[T],\, a\in\A \qquad
N_{t,a} =0 \quad\Rightarrow\quad
    \hat{\mu}^t_a \geq \tfrac13.
\]

\item[(c)] Reward estimates $\hat{\mu}_a^t$ depend only on the anonymized subhistory $\AnonSubH{S_t}$. That is, reward estimates depend only on the observations, not on the rounds in which these observations are collected.

\item[(c)] Reward estimates are consistent over time, but allow some agent heterogeneity. Formally, each agent $t$ forms its estimates according to a \emph{reward estimate function}: a function $f_t$ from anonymous subhistories to $[0,1]^K$, so that the estimate vector
        $(\hat{\mu}_{t,a}:\, a\in\A)$
    equals $f_t(\AnonSubH{S_t})$. And this function is drawn independently from some fixed distribution over reward estimate functions.
\end{itemize}
\end{assumption}

@@

\xhdr{Connection to multi-armed bandits.}
The special case when each message $m_t$ is an arm, and the $t$-th agent always chooses this arm, corresponds to a standard multi-armed bandit problem with IID rewards. Following the literature on bandits, we define the \emph{gap parameter} $\Delta$ as the difference between the largest and second largest mean rewards.%
\footnote{Formally, the second-largest mean reward is 
    $\max_{a\in\A:\mu(a)<\mu^*} \mu(a)$,
where $\mu^* = \max_{a\in\A} \mu(a)$.
}
The gap parameter is not known to the principal (in our problem), or to the algorithm (in the bandit problem). However, it is essential for regret bounds.

\ascomment{We'll need to state/discuss optimal regret bounds somewhere, probably here.}

>>>>

We study the multi-armed bandit problem in a social learning context, in which a principal faces a sequence of $T$ myopic agents. There is a set $\A$ of $K$ possible actions, a.k.a. \emph{arms}. At each round $t\in [T]$, a new agent $t$ arrives, receives a message $m_t$ from the principal, chooses an arm $a_t\in \A$, and collects a reward $r_t\in \{0,1\}$ that is immediately observed by the principal. The reward from pulling an arm $a\in \A$ is drawn independently from Bernoulli distribution $\cD_a$ with an unknown mean $\mu_a$. An agent does not observe anything from the previous rounds, other than the message $m_t$. The problem instance is defined by (known) parameters $K,T$ and the (unknown) tuple of mean rewards, $(\mu_a:\,a\in\A)$. We are interested in \emph{regret}, defined as
\begin{align}
\textstyle
  \reg(T)
  %= \reg(a_1, \ldots , a_T)
  = T \max_{a\in \A} \mu_a -
  \sum_{t\in [T]} \E[\mu_{a_t}].
\end{align}
The principal chooses messages $m_t$ according to an online algorithm called \emph{disclosure policy}, with a goal to minimize regret. We assume that mean rewards are bounded away from $0$ and $1$, to ensure sufficient entropy in rewards. For concreteness, we posit that
    $\mu_a\in [\tfrac13,\tfrac23]$.

\xhdr{Unbiased subhistories.}
The \emph{subhistory} for a subset of rounds $S\subset [T]$ is defined as 
\begin{align} \label{eq:model-subhistory}
    \SubH{S} = \left\{\; (s,a_s,r_s):\;s\in S \;\right\}.
\end{align}
Accordingly, $\SubH{[t-1]}$ is called the \emph{full history} at time $t$. 
The \emph{outcome} for agent $t$ is the tuple $(t,a_t,r_t)$.

@@

Further, we focus on disclosure policies that are \emph{transitive}, in the following sense:
\[ t'\in S_t \Rightarrow S_{t'}\subset S_t
    \qquad \text{for all rounds $t,t'\in [T]$}. \]
In words, if an agent observes the outcome for some previous agent, then she observes the entire message revealed to that agent.

A transitive unbiased-history policy can be represented as an undirected graph, where nodes correspond to rounds, and any two rounds $t'<t$ are connected if and only if $t'\in S_t$ and there is no intermediate round $t''$ with
    $t'\in S_{t''}$ and $t''\in S_t$.
This graph is henceforth called the \emph{information flow graph} of policy $\pi$, or \emph{info-graph} for short. We assume that it is common knowledge.

\xhdr{Agents' behavior.} Let us define agents' behavior in response to an unbiased-history policy. We posit that each agent $t$ uses its observed subhistory $m_t$ to form a reward estimate $\hat{\mu}_{t,a} \in [0,1]$ for each arm $a\in \A$, and chooses an arm with a maximal estimator. (Ties are broken according to an arbitrary rule that is the same for all agents.) The basic model is that $\hat{\mu}_{t,a}$ is the sample average for arm $a$ over the subhistory $m_t$, as long as it includes at least one sample for $a$; else, $\hat{\mu}_{t,a}\geq \tfrac13$.

We allow a much more permissive model, characterized by the following assumption:

@@

\begin{assumption}\label{ass:embehave}
Reward estimates are close to empirical averages. Let $N_{t,a}$ and $\bar{\mu}_{t,a}$ denote the number of pulls and the empirical mean reward of arm $a$ in subhistory $m_t$. Then for some absolute constant $N_0\in \N$ and $C_0=\tfrac{1}{16}$, and for all agents $t\in [T]$ and arms $a\in\A$ it holds that
\[
\text{if}\; N_{t,a} \geq N_0 
\quad\text{then}\quad
    \left|\hat{\mu}^t_a - \bar{\mu}^t_a \right| <
		\frac{C_0}{\sqrt{N_{t,a}}}.
\]
Also, 
    $\hat{\mu}^t_a\geq\tfrac13$ if $N_{t,a}=0$.
(NB: we make no assumption if $1\leq N_{t,a}<N_0$.)
\end{assumption}

\OMIT{ %%%%%%%%
\begin{assumption}
Reward estimates are consistent over time, but allow some agent heterogeneity. Formally, each agent $t$ forms its estimates according to a \emph{reward estimate function}: a function $f_t$ from subhistories to $[0,1]^K$, so that the estimate vector
        $(\hat{\mu}_{t,a}:\, a\in\A)$
    equals $f_t(m_t)$. This function is drawn independently from some fixed distribution over reward estimate functions.
\end{assumption}
} %%%%%%%

@@

\xhdr{Connection to multi-armed bandits.}
The special case when each message $m_t$ is an arm, and the $t$-th agent always chooses this arm, corresponds to a standard multi-armed bandit problem with IID rewards. Thus, regret in our problem can be directly compared to regret in the bandit problem with the same mean rewards $(\mu_a:\,a\in\A)$. Following the literature on bandits, we define the \emph{gap parameter} $\Delta$ as the difference between the largest and second largest mean rewards.%
\footnote{Formally, the second-largest mean reward is
    $\max_{a\in\A:\mu(a)<\mu^*} \mu(a)$,
where $\mu^* = \max_{a\in\A} \mu(a)$.
}
The gap parameter is not known to the principal (in our problem), or to the algorithm (in the bandit problem). Optimal regret rates for bandits with IID rewards are as follows \cite{bandits-ucb1,bandits-exp3,Lai-Robbins-85}: 
\begin{align}\label{eq:model-OptRegret}
\reg(T) \leq O\left(\min\left( 
    \sqrt{KT\log T)},\; \tfrac{1}{\Delta} \log T
    \right)\right).
\end{align}
This regret bound can only be achieved using \emph{adaptive exploration}: \ie when exploration schedule is adapted to the observations. A simple example of \emph{non}-adaptive exploration is the \emph{explore-then-exploit} algorithm which samples arms uniformly at random for the first $N$ rounds, for some pre-set number $N$, then chooses one arm and sticks with it till the end. More generally, 
\emph{exploration-separating} algorithms have a property that in each round $t$, either the choice of an arm does not depend on the observations so far, or the reward collected in this round is not used in the subsequent rounds. Any such algorithm suffers from $\Omega(T^{2/3})$ regret in the worst case.%
\footnote{The first explicit reference we are aware of is \cite{MechMAB-ec09,DevanurK09}, but this fact has been known in the community for much longer.}

\xhdr{Preliminaries.}
Throughout the paper, we will use the standard concentration and anti-concentration inequalities: respectively, Chernoff Bounds and Berry-Esseen Theorem. The former states that that a sum of independent random variables converges to its expectation quickly. The latter states that the CDF of an appropriately scaled average of IID random variables converges to the CDF of the standard normal distribution pointwise. In particular, the average strays far enough from its expectation with some guaranteed probability. The theorem statements are moved to Appendix~\ref{sec:prelim}.
